### 2022-01-03 slides/IBC2020/oral/v2/pres.tex

\documentclass{beamer}

\mode<presentation>
{
  \usetheme{Goettingen}      
  \usecolortheme{default}
  \usefonttheme{default}
  \setbeamertemplate{navigation symbols}{}
  \setbeamertemplate{caption}[numbered]
}

\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage{amsmath}
\usepackage{dsfont}
\usepackage{booktabs,multirow,array}
\usepackage{algorithm}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{graphicx}
\usepackage{url}
\usepackage{xcolor}
\usepackage{moresize}
\usepackage{lmodern}
\usepackage{mathrsfs}

\usepackage[numbers]{natbib}
\renewcommand\bibsection{\section[]{\refname}}

\definecolor{blue_pres}{RGB}{51,50,178}
\definecolor{darkpastelgreen}{rgb}{0.01, 0.75, 0.24}

\hypersetup{
    colorlinks,
    citecolor=orange,
    linkcolor=blue_pres,
    backref=true,
    urlcolor=darkpastelgreen
}

\makeatletter
\let\@@magyar@captionfix\relax
\makeatother

\DeclareCaptionFormat{myformat}{#3}
\captionsetup[algorithm]{format=myformat}
\usepackage{algpseudocode}

%%%%%%%%%%%%%%%%%%%%
%%% New commands %%%
%%%%%%%%%%%%%%%%%%%%

\DeclareMathOperator{\TV}{TV}
\DeclareMathOperator{\bina}{bina}
\DeclareMathOperator{\argmin}{argmin}

\newcommand{\dd}{\mathrm{d}}
\newcommand{\bC}{\mathbb C}
\newcommand{\bD}{\mathbb D}
\newcommand{\bP}{\mathbb P}
\newcommand{\bX}{\boldsymbol X}
\newcommand{\XB}{{\boldsymbol X}^B}
\newcommand{\E}{\mathbb E}
\newcommand{\N}{\mathbb N}
\newcommand{\bigO}{\mathcal{O}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\D}{\mathbb{D}}
\newcommand{\R}{\mathbb R}
\newcommand{\cA}{\mathcal A}
\newcommand{\cB}{\mathcal B}
\newcommand{\cH}{\mathcal H}
\newcommand{\cS}{\mathcal S}
\newcommand{\cM}{\mathcal M}
\newcommand{\cC}{\mathcal C}
\newcommand{\cG}{\mathcal G}
\newcommand{\cN}{\mathcal N}
\newcommand{\cU}{\mathcal U}
\renewcommand{\P}{\mathds{P}}
\newcommand{\Gb}{\bar G}
\newcommand{\Fb}{\bar F}
\newcommand{\Ub}{\bar U}
\newcommand{\by}{\boldsymbol y}
\newcommand{\bz}{\boldsymbol z}
\newcommand{\bdelta}{\boldsymbol \delta}
\newcommand{\bSigma}{\textbf{$\Sigma$}}
\newcommand{\norm}[1]{\|#1\|}
\newcommand{\pluseq}{\mathrel{+}=}
\newcommand{\minuseq}{\mathrel{-}=}
\newcommand{\bcdot}{\raisebox{-0.25ex}{\scalebox{1.2}{$\cdot$}}}
\newcommand{\ind}[1]{\mathds{1}_{#1}}
\newcommand\independent{\protect\mathpalette{\protect\independenT}{\perp}} 
\newcommand{\cD}{\mathcal D}
\newcommand{\cQ}{\mathcal Q}
\newcommand{\cR}{\mathcal R}
\renewcommand{\P}{\mathds P}

\definecolor{blue_pres}{RGB}{51,50,178}

\newcommand\Algphase[1]{%
\vspace*{-.7\baselineskip}\Statex\hspace*{\dimexpr-\algorithmicindent-2pt\relax}\rule{\textwidth}{0.4pt}%
\vspace*{-.4\baselineskip}\Statex\hspace*{\algorithmicindent}\textbf{#1}%
\vspace*{-.7\baselineskip}\Statex\hspace*{\dimexpr-\algorithmicindent-2pt\relax}\rule{\textwidth}{0.4pt}%
}
\newcommand*\rot{\rotatebox{90}}
\newcommand*\OK{\ding{51}}

\def\independenT#1#2{\mathrel{\rlap{$#1#2$}\mkern2mu{#1#2}}}
\newtheorem{hyp}{Hypothesis}

\title[IBC 2020 \\ Lights \\ \insertframenumber/\inserttotalframenumber \vspace{-1cm}]{}

\institute[shortinst]{
  \vspace{-2.2cm} \\
  \includegraphics[width=2.5cm]{figures/SU} \hspace{1cm}
  \includegraphics[width=2.2cm]{figures/inserm} \hspace{1cm}
  \includegraphics[width=1.7cm]{figures/Califrais_logo} \\
  \vspace{.6cm}
  \textcolor{blue_pres}{\Large IBC 2020 Online Learning Series}\\
  \vspace{.5cm}
  \large
  \textbf{Simon Bussy}\\
  \vspace{.25cm}
  \scriptsize
  \href{mailto:simon.bussy@gmail.com}{\texttt{simon.bussy@gmail.com}}\\
  \vspace{.4cm}
  \small
  \textbf{Lights: a generalized joint model for high-dimensional multivariate longitudinal data and censored durations} \\
  \vspace{.3cm}
  S. Bussy $^{(1,2)}$, A. Barbieri $^{(3)}$, S. Zohar $^{(1)}$, A.S. Jannot $^{(1,4)}$\\

  \vspace{.4cm}

  \tiny

  $^{(1)}$INSERM, UMRS 1138, Centre de Recherche des Cordeliers, Paris, France. \\
  $^{(2)}$Califrais' Machine Learning Lab, Paris, France. \\
  $^{(3)}$INSERM, UMR 1219, Bordeaux Population Health Research Center, Univ. Bordeaux, France. \\
  $^{(4)}$Biomedical Informatics and Public Health Department, EGPH, APHP, Paris, France. 
}

\date{}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\section{Introduction}

\begin{frame}[noframenumbering]
\Large \centering
\textcolor{blue_pres}{I.} Introduction
\end{frame}

\subsection{Overview}

\begin{frame}{Overview}

\begin{itemize}
  \item<1-> Deal with the problem of joint modeling of longitudinal data and censored durations
  \item<2-> Large number of both longitudinal and time-independent features are available
  \item<3-> Flexibility in modeling the dependency between the longitudinal features and the event time with appropriate penalties
  \item<4-> Inference achieved using an efficient and novel Quasi-Newton Monte Carlo Expectation Maximization algorithm
\end{itemize}

\end{frame}

\subsection{Use cases}

\begin{frame}{Use cases}

\small
\begin{itemize}
  \item<1-> Predict the risk for an event of interest to occur quickly, taking into account simultaneously a huge number of longitudinal signals in a high-dimensional context
  \item<2-> Provides powerful interpretability by automatically pinpointing significant time-dependent and time-independent features
\end{itemize}

\begin{block}<3->{Real-time decision support}
\begin{itemize}
  \item<3-> Medical context $\rightarrow$ event of interest: survival time, re-hospitalization, relapse or disease progression ; longitudinal data: biomarkers or vital parameters measurements
  \item<4-> Customer's satisfaction monitoring context $\rightarrow$ event of interest: time when a client churns ; longitudinal data: the client's activity recorded from account opening throughout the duration of the business relationship
\end{itemize}
\end{block}

\end{frame}

\subsection{Framework}

\begin{frame}{High-dimensional framework}

\small
\begin{itemize}
  \item<1-> Survival analysis
  \begin{center}
  $T = T^\star \wedge C \quad \text{and} \quad \Delta = \ind{\{T^\star \leq C\}}$
  \end{center}
  \item<2-> Time-independent features $X \in \R^p$ with $p \gg n$
  \item<3-> $L$ longitudinal outcomes such that $L \gg n$ and \[Y(t) = \big(Y^1(t), \ldots, Y^L(t) \big)^\top \in \R^L\]
  \item<4-> Heterogeneity of the population: latent subgroups \[G \in \{ 0, \ldots, K-1\}\]
  \item<5-> Softmax link function for the latent class membership probability given time-independent features
  \begin{center}
  $\pi_{\xi_k}(x) = \P[G=k|X=x] = \dfrac{e^{x^\top\xi_k}}{\sum_{k=0}^{K-1}e^{x^\top\xi_k}}$
  \end{center}
\end{itemize}

\end{frame}

\section{Model}

\begin{frame}[noframenumbering]
\Large \centering
\textcolor{blue_pres}{II.} Model
\end{frame}

\subsection{Submodels}

\begin{frame}{Submodels}

\small
\begin{block}<1->{Group-specific marker trajectories}
\begin{itemize}
  \item<1-> \footnotesize $h_l\big(\E[Y^l(t)|b^l, G=k]\big) = m_k^l(t) = u^l(t)^\top\beta_k^l + v^l(t)^\top b^l$\\
  with fixed effect parameters $\beta_k^l \in \R^{q_l}$ and subject-and-longitudinal outcome specific random effects $b^l \in \R^{r_l} \sim \cN(0, D_{ll})$
  \item<2-> $\text{Cov}[b^l,b^{l'}] = D_{ll'}$
  \only<2>{ and
  \[ D = 
  \begin{bmatrix}
    D_{11} & \cdots & D_{1L}\\
    \vdots &  \ddots & \vdots \\
    D_{1L}^\top & \cdots & D_{LL}
  \end{bmatrix}
  \]
  the global variance-covariance matrix}

\end{itemize}
\end{block}

\begin{block}<3->{Group-specific risk of event}
\begin{itemize}
  \item<3-> \footnotesize $\lambda(t|G = k) = \lambda_0(t) \exp \Big\{ x^\top \gamma_{k,0} + \sum_{l=1}^L \sum_{a=1}^\cA {\gamma_{k,a}^l}^\top \varphi_a(t, \beta_k^l, b^l) \Big\}$
  \item<4-> Functionals $(\varphi_a)_{a\in \cA}$
\begin{table}[htb]
\centering
\resizebox{.9\textwidth}{!}{%
\begin{tabular}{cccc}
\toprule
Description & $\varphi_a(t, \beta_k^l, b^l)$ & $\frac{\partial \varphi_a(t, \beta_k^l, b^l)}{\partial \beta_k^l}$  & Reference \\
\midrule
Linear predictor & $m_k^l(t)$ & $u^l(t)$ & \citet{chi2006joint} \\ [.15cm]
Random effects & $b^l$ & $\mathbf{0}_{q_l}$ & \citet{hatfield2011joint} \\ 
Time-dependent slope & $\dfrac{\dd}{\dd t} m_k^l(t)$ & $\dfrac{\dd}{\dd t} u^l(t)$ & \citet{rizopoulos2011bayesian} \\ [.3cm]
Cumulative effect & $\int_0^t m_k^l(s) \dd s$ & $\int_0^t u^l(s) \dd s$ & \citet{andrinopoulou2017combined} \\ [.1cm]
 \bottomrule
\end{tabular}}
\end{table}

\end{itemize}
\end{block}

\end{frame}

\subsection{Likelihood}

\begin{frame}{Likelihood}

\footnotesize
\begin{itemize}
  \item<1-> $\cD_n = \big\{ (x_1, y_1^1, \ldots, y_1^L, t_1, \delta_1), \ldots, (x_n, y_n^1, \ldots, y_n^L, t_n, \delta_n) \big\}$ with $y_i^l=(y_{i1}^l, \ldots, y_{in_i^l}^l)^\top \in \R^{n_i^l} \text{ and } y_{ij}^l=Y_i^l(t_{ij}^l)$
  \item<2-> $y_i = ({y_i^1}^\top \cdots {y_i^L}^\top)^\top \in \R^{n_i}$ with $n_i = \sum_{l=1}^L n_i^l$
  \item<3-> $f(y_i|b_i, G_i=k) = \exp \big\{(y_i \odot \Phi_i)^\top M_{ik} - c_\phi(M_{ik}) + d_\phi(y_i) \big\}$ with $\Phi_i = (\phi_1^{-1} {\textbf{1}_{n_i^1}}^\top \cdots \phi_L^{-1} {\textbf{1}_{n_i^L}}^\top)^\top \in \R^{n_i}$
  \item<4-> \scriptsize $\theta = \big(\xi_0^\top \cdots \xi_{K-1}^\top, \beta_0^\top \cdots \beta_{K-1}^\top, \phi^\top, \text{vech}(D), \lambda_0(t), \gamma_0^\top \cdots \gamma_{K-1}^\top\big) \in \R^\vartheta$ 
  \item<5-> Survival part: 
  \begin{align*}f(t_i, \delta_i| b_i, G_i = k ; \theta) = &\big[\lambda(t_i|\cM_k(t_i), G_i = k)\big]^{\delta_i} \\ & \times \exp \Big\{-\int_0^{t_i} \lambda(s|\cM_k(s), G_i = k) \dd s \Big\}
  \end{align*}
  \item<6-> Then, the likelihood writes
  \begin{align*}\ell_n(\theta) = n^{-1} \sum_{i=1}^n \log \int_{\R^r} \sum_{k=0}^{K-1} &\pi_{\xi_k}(x_i) f(t_i, \delta_i| b_i, G_i = k ; \theta) \\ & \times f(y_i | b_i, G_i = k ; \theta) f(b_i ; \theta) \dd b_i
  \end{align*}
\end{itemize}

\end{frame}

\section{Inference}

\begin{frame}[noframenumbering]
\Large \centering
\textcolor{blue_pres}{III.} Inference
\end{frame}

\subsection{Penalization}

\begin{frame}{Penalization}

\small

\begin{itemize}
  \item<1-> Penalized objective
  \begin{equation*}
  \ell_n^\text{pen}(\theta) = - \ell_n(\theta) + \sum_{k=0}^{K-1} \zeta_{1,k} \norm{\xi_k}_{\text{en}, \eta} + \zeta_{2,k} \norm{\gamma_k}_{\text{sg} l_1, \tilde{\eta}} + \zeta_{3,k} \norm{\beta_k}_{\text{sg} l_1, \tilde{\eta}}
  \end{equation*}
  with the elasticnet penalty \[ \norm{z}_{\text{en}, \eta} = (1-\eta)\norm{z}_1 + \dfrac\eta2 \norm{z}_2^2 \] and the sparse group lasso penalty \[ \norm{z}_{\text{sg} l_1, \tilde{\eta}} = (1-\tilde{\eta})\norm{z}_1 + \tilde{\eta} \sum_{l=1}^L\norm{z^l}_2 \]
  \item<2-> Resulting optimization problem \[\hat \theta \in \argmin_{\theta \in \R^\vartheta} \ell_n^\text{pen}(\theta)\]
\end{itemize}

\end{frame}

\subsection{MCQNEM}

\begin{frame}{MCQNEM algorithm}

\small

\begin{itemize}
  \item<1-> $\ell_n^\text{comp}(\theta) = \ell_n^\text{comp}(\theta ; \cD_n, \textbf{\textit{b}}, \textbf{\textit{G}})$
\end{itemize}

\begin{block}<2->{Monte Carlo E-step}
\begin{itemize}
  \item<2-> $\cQ_n(\theta, \theta^{(w)}) = \E_{\theta^{(w)}}[\ell_n^\text{comp}(\theta) | \cD_n]$
\end{itemize}
\end{block}

\begin{block}<3->{Quasi-Newton M-step}
\begin{itemize}
  \item<3-> \tiny $\theta^{(w+1)} \in \argmin_{\theta} \cQ_n(\theta, \theta^{(w)}) + \sum_{k=0}^{K-1} \zeta_{1,k} \norm{\xi_k}_{\text{en}, \eta} + \zeta_{2,k} \norm{\gamma_k}_{\text{sg} l_1, \tilde{\eta}} + \zeta_{3,k} \norm{\beta_k}_{\text{sg} l_1, \tilde{\eta}}$
\end{itemize}
\end{block}

\begin{itemize}
\item<4-> Predictive marker \[\hat \cR_{ik} = \dfrac{\pi_{\hat \xi_k}(x_i) \hat f(t^{max}_i, y_i | b_i, G_i=k ; \hat \theta)}{\sum_{k=0}^{K-1} \pi_{\hat \xi_k}(x_i) \hat f(t^{max}_i, y_i | b_i, G_i=k ; \hat \theta)},\] which is an estimate of \[\P_\theta[G_i=k | T^\star_i > t^{max}_i, y_i]\]
\end{itemize}

\end{frame}

\section{Conclusion}

\begin{frame}[noframenumbering]
\Large \centering
\textcolor{blue_pres}{V.} Conclusion
\end{frame}

\begin{frame}{Conclusion}

\begin{itemize}
  \item<1-> Prognostic method called lights introduced to deal with the problem of joint modeling of longitudinal data and censored durations, where a large number of longitudinal features are available
  \item<2-> Penalization of the likelihood in order to perform feature selection and to prevent overfitting
  \item<3-> New efficient estimation algorithm (QNMCEM) has been derived
  \item<4-> Automatically determines significant prognostic longitudinal features
\end{itemize}

\begin{block}<4->{\texttt{Python} 3 package}
\begin{itemize}
  \item<4-> Available at \small{\url{https://github.com/Califrais/lights}}
  \item<5-> Applications of the model available soon on an \texttt{arXiv} paper.
\end{itemize}
\end{block}

\end{frame}

\begin{frame}[noframenumbering]
\Large \centering
\textcolor{blue_pres}{} Thank you!
\end{frame}

\section{References}
\frame{
  \small
  \frametitle{References}
  \bibliographystyle{plainnat}
  \bibliography{refs}
}

\end{document}